---
title: "My first R Markdown!"
author: "Sam Csik"
date: "6/8/2021"
output: 
  html_document:
    toc: true
    code_folding: hide
    theme: yeti
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

## Agenda

  1. create/knit an R Markdown file
  2. embed outputs into our knitted doc
  3. discuss ways R Markdown may be useful  
  ----------------------------------------
  4. organize! --> YAML metadata header
      - table of contents
      - collapsible code chunks
      - theme
      - tabs*
  5. improve accessibility
      - alt text
  6. publish as a website --> GitHub Pages

```{r, message = FALSE, warning = FALSE, echo = FALSE} 
library(palmerpenguins) 
library(tidyverse)
library(DT)
library(gt)
```

## Introduction

Here, I'm providing some background information about penguins, and why it's important for us to take different types of morphological measurements. Here are the methods I used. Etc. 

I can also hyperlink to other resources. For example, see a more complete description of the `palmerpenguins` package and data [here](https://allisonhorst.github.io/palmerpenguins/index.html).

## Explore the raw data

```{r, warning = FALSE, message = FALSE}
##############################
# generate an interactive table (using the `DT` package) 
##############################

table1 <- datatable(penguins, 
                    class = 'cell-border stripe', 
                    colnames = c("Species", "Island", "Bill length (mm)",
                                 "Bill depth (mm)", "Flipper length (mm)",
                                 "Body mass (g)", "Sex", "Year"),
                    caption = htmltools::tags$caption( 
                      style = 'caption-side: top; text-align: left;', 
                      htmltools::em('Table 1: Morphological data for male and female Adelie, Chinstrap, & Gentoo penguins across three islands in the Palmer Archipelago, Antarctica.')),
                    filter = 'top', 
                    options = list(pageLength = 5, autoWidth = TRUE)) 

##############################
# display table in knitted doc
##############################

table1
```

## Penguin sizes {.tabset .tabset-fade .tabset-pills}

### Bill length ~ body mass
```{r, warning = FALSE, fig.cap = "**Fig. 1:** Penguin bill length ~ body mass for Adelie, Chinstrap, & Gentoo Penguins", fig.alt = "A scatterplot of body mass (in grams) by bill length (in millimeters) for Adelie, Chinstrap, and Gentoo penguins, where we show that the penguins with greater body mass tend to have larger bills."}
ggplot(data = penguins, aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(aes(shape = species), alpha = 0.8) +
  scale_color_manual(values = c("darkorange","purple","darkcyan")) +
  labs(x = "Body mass (g)", y = "Bill length (mm)") +
  theme_minimal()
```

### Body size summary statistics
```{r, warning = FALSE}
##############################
# summarize df
##############################

penguin_summary <- penguins %>% 
  drop_na() %>% 
  group_by(species) %>% 
  summarize(
    mean_body_mass = round(mean(body_mass_g), 2), 
    sd_body_mass = round(sd(body_mass_g), 2),
    mean_bill_length = round(mean(bill_length_mm), 2),
    sd_bill_length = round(sd(bill_length_mm), 2)
  )

##############################
# create table with DT
##############################

table2 <- penguin_summary %>% 
  gt() %>% 
  tab_header(title = "Mean body mass and bill length measurements (±SD) for Adelie, Chinstrap, & Gentoo penguins") %>% 
  cols_label(species = "Species",
             mean_body_mass = "Mean body mass (g)",
             sd_body_mass = "SD body mass (g)",
             mean_bill_length = "Mean bill length (mm)",
             sd_bill_length = "SD bill length (mm)") %>% 
  cols_align("center") %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_title(groups = "title")
  ) %>% 
  tab_style(
    style = cell_text(color = "darkorange"),
    locations = cells_body(
      rows = species == "Adelie")
  ) %>% 
    tab_style(
    style = cell_text(color = "purple"),
    locations = cells_body(
      rows = species == "Chinstrap")
  ) %>%
    tab_style(
    style = cell_text(color = "darkcyan"),
    locations = cells_body(
      rows = species == "Gentoo")
  )

##############################
# display table in knitted doc
##############################

table2
```

